{\plant}: A package for modelling forest trait ecology \& evolution

### 2015-08-27 ms.tex
Minor comments in a few places on changes

Broadly, the {\plant} model falls into a class we refer to Trait- Size- and
Patch- structured Models (\textsc{TSPMs}). \textsc{TSPMs} are direct
descendants of the ``gap'' models developed in 1980s
\citep[e.g.][]{Shugart-1980, Huston-1987, Kohyama-1993}; other modern examples
include \textsc{ED} \citep{Moorcroft-2001}, \textsc{LPJ- GUESS}
\citep{Smith-2014} and \textsc{LPJml-FIT} \citep{Sakschewski-2015}. Common to
\textsc{TSPMs} are that they explicitly model size-structured competition
among individual plants within a metapopulation of ``patches'', where
individuals can also differ in their traits and thus demographic behaviour.
\textsc {TSPMs} also typically ignore the spatial structure of plants within
patches and the spatial structure of patches relative to one another; instead
focussing computational effort on resolving the dynamics of size-structured
competition for light and successional turnover. Differences among
\textsc{TSPMs} arise from both the core physiological models used (describing
how the demography of individual plants responds to resource availability),
and the numerical techniques applied to scaling the physiological model up to
estimate emergent behaviours of individuals, patches and vegetation.

@@

{\plant} is an individual-based model, meaning the dynamics of the
population arise from rules specifying how individuals grow and interact. 
Being driven by traits, the model can be extended to potentially very many species. The
core rules in {\plant} are about the short-term physiological
functioning of an individual plant and how these are influenced by its
traits, size and light environment.  While {\plant}
includes a core physiological model, we use it here only as an example to
demonstrate the different ways in which any given physiological model
could be analysed. To be clear, it is not the purpose
of this paper to justify the particular physiological model used, to provide detailed
comparisons to data, or necessarily promote this physiological model over
those in other \textsc{TSPMs}. (These are all topics worthy of
substantially more attention than can be afforded here.) Moreover, an important
feature of {\plant} is that it allows for users to add their own
physiological models.

@@

While we have implemented one physiological model (\texttt{FFW16}), the
package is designed to allow arbitrary additional physiologies to be
introduced. Users are able to modify the physiological model in three
different ways. First, one can vary the parameters of the \texttt{FFW16} model
directly (see Table XXX for list, for examples see Appendix
\ref{sec:parameters}). Second, we allow for changes in key parameters to
trigger changes in other parameters (i.e. a hyper-parametrisation), enabling
straightforward modelling of trade-offs. For example, the trait \textsc{lma}
(leaf mass per unit leaf area) is used to estimate the rate of leaf turnover
($k_l$), based on an observed scaling relationship spanning across diverse
vegetation and plant functional types \citep{Wright-2004}:
$$k_l = \beta_{kl1} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{kl2}},$$ 
where $k_l$ and $\phi$ (lma) are both parameters of the core physiological
model, and $\beta_{kl1}, \beta_{kl2}, \phi_0$ are hyper-parameters. Such
linkages are defined within a user-supplied hyper-parameterisation function
(see Appendix \ref{sec:FFW16} for details). Finally, rather than varying the
parameters of the existing equations in the \texttt{FFW16} model, the
equations themselves can be changed and the higher level functionality of the
package can still be used.  The latter corresponds to defining a new
physiological model. As an example, in the \texttt{FFW16b} physiology, we have modified
the function determining allocation to reproduction. The specifications
for writing a new physiological model are outlined further in Appendix 
XXX.

@@

When modelling a patch of competing, the focus is on changes to the size-
density-distribution \(N(x, H, a)\) with patch age \(a\), i.e. the density of
individuals with traits \(x\) and height \(H\) within a patch of age \(a\).
Density is measured as individuals per unit height and per unit ground area.
Modelling the development of  \(N(x, H, a)\) requires that both the initial size
distribution and the inflow of new recruits be specified. We have primarily
been interested in dynamics of a patch recovering from disturbance and have
therefore started with an empty patch and constant flow of seeds from a global
seed rain (Fig. \ref{fig:schematic}b).

@@

@@

Other \textsc{TSPMs} also exist for modelling of vegetation, like \textsc{ED}
\citep[ver 1 and 2][]{Moorcroft-2001, Medvigy-2009}, \textsc {LPJ-GUESS}
\citep{Smith-2014} and a\textsc{LPJml-FIT}  \citep{Sakschewski-2015}. These
models were designed for simulation of bio-geochemical cycles and vegetation
structure across broad geographic areas; as such their architectures are
optimised for these use-cases. By contrast, {\plant} was  designed so that the
physiological rules driving the dynamics of the system could be easily
manipulated and emergent demographic behaviours (including fitness) studied in
detail. These different use-cases underpin important design differences among
the models. Whereas the primary purpose of these other models is to scale a
given physiology up to predict  broad-scale changes in the terrestrial
biosphere, {\plant} allows users to  interact with the model at multiple
levels of ecological organisation and  study the dynamics in detail. With any
given physiology, it is possible to  grow plants in isolation, in patches of
competing plants and in disturbance- driven metapopulations. {\plant} also
allows for greater focus on population dynamics, reporting not only vegetation
properties but also individual level densities, growth rates and fitness.
{\plant} thus aims to bridge the gap between existing large-scale coupled
simulation models like \textsc{ED} and the extremely simple models that are
still widely  used in theoretical ecology.

>>>>

% TODO[RGF]: Is this new acronym needed?  It's not really the place of
% this contribution to define a new class of models, and given that
% this all comes from a response to a reviewer, IMO it bends the paper
% to be too much like a response to the reviewer.
Broadly, the {\plant} model falls into a class we refer to Trait- Size- and
Patch- structured Models (\textsc{TSPMs}). \textsc{TSPMs} are direct
descendants of the ``gap'' models developed in 1980s
\citep[e.g.][]{Shugart-1980, Huston-1987, Kohyama-1993}; other modern examples
include \textsc{ED} \citep{Moorcroft-2001}, \textsc{LPJ- GUESS}
\citep{Smith-2014} and \textsc{LPJml-FIT} \citep{Sakschewski-2015}. Common to
\textsc{TSPMs} are that they explicitly model size-structured competition
among individual plants within a metapopulation of ``patches'', where
individuals can also differ in their traits and thus demographic behaviour.
\textsc {TSPMs} also typically ignore the spatial structure of plants within
patches and the spatial structure of patches relative to one another; instead
focussing computational effort on resolving the dynamics of size-structured
competition for light and successional turnover. Differences among
\textsc{TSPMs} arise from both the core physiological models used (describing
how the demography of individual plants responds to resource availability),
and the numerical techniques applied to scaling the physiological model up to
estimate emergent behaviours of individuals, patches and vegetation.

@@

{\plant} is an individual-based model, meaning the dynamics of the
population arise from rules specifying how individuals grow and interact. 
Being driven by traits, the model can be extended to potentially very many species. The
core rules in {\plant} are about the short-term physiological
functioning of an individual plant and how these are influenced by its
traits, size and light environment.  While {\plant}
includes a core physiological model, we use it here primarily as an example to
demonstrate the different ways in which any given physiological model
could be analysed. To be clear, it is not the purpose
of this paper to justify the particular physiological model used, to provide detailed
comparisons to data, or necessarily promote this physiological model over
those in other \textsc{TSPMs}. (These are all topics worthy of
substantially more attention than can be afforded here.) Moreover, an important
feature of {\plant} is that it allows for users to add their own
physiological models.

@@

%% TODO[RGF]: I think that this needs softening.  Users can't really
%% add new models; that has to be done by *developers*, really
%% (because of the design the models are compiled into the package.
While we have implemented one physiological model (\texttt{FFW16}), the
package is designed to allow arbitrary additional physiologies to be
introduced. Users are able to modify the physiological model in three
different ways. First, one can vary the parameters of the \texttt{FFW16} model
directly (see Table XXX for list, for examples see Appendix
\ref{sec:parameters}). Second, we allow for changes in key parameters to
trigger changes in other parameters (i.e. a hyper-parametrisation), enabling
straightforward modelling of trade-offs. For example, the trait \textsc{lma}
(leaf mass per unit leaf area) is used to estimate the rate of leaf turnover
($k_l$), based on an observed scaling relationship spanning across diverse
vegetation and plant functional types \citep{Wright-2004}:
$$k_l = \beta_{kl1} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{kl2}},$$ 
where $k_l$ and $\phi$ (lma) are both parameters of the core physiological
model, and $\beta_{kl1}, \beta_{kl2}, \phi_0$ are hyper-parameters. Such
linkages are defined within a user-supplied hyper-parameterisation function
(see Appendix \ref{sec:FFW16} for details). Finally, rather than varying the
parameters of the existing equations in the \texttt{FFW16} model, the
equations themselves can be changed and the higher level functionality of the
package can still be used.  The latter corresponds to defining a new
physiological model. As an example, in the \texttt{FFW16b} physiology, we have modified
the function determining allocation to reproduction.

@@

When modelling a patch of competing plants, the focus is on changes to the size-
density-distribution \(N(x, H, a)\) with patch age \(a\), i.e. the density of
individuals with traits \(x\) and height \(H\) within a patch of age \(a\).
Density is measured as individuals per unit height and per unit ground area.
Modelling the development of  \(N(x, H, a)\) requires that both the initial size
distribution and the inflow of new recruits be specified. We have primarily
been interested in dynamics of a patch recovering from disturbance and have
therefore started with an empty patch and constant flow of seeds from a global
seed rain (Fig. \ref{fig:schematic}b).

@@

% TODO[RGF]: I think that this section should be reduced considerably,
% ideally into a single paragraph.  The first three "paragraphs"
% should be largely retained, shrinking the third.

@@

Other \textsc{TSPMs} also exist for modelling of vegetation, like \textsc{ED}
\citep[ver 1 and 2][]{Moorcroft-2001, Medvigy-2009}, \textsc {LPJ-GUESS}
\citep{Smith-2014} and a\textsc{LPJml-FIT}  \citep{Sakschewski-2015}. These
models were designed for simulation of bio-geochemical cycles and vegetation
structure across broad geographic areas; as such their architectures are
optimised for these use-cases. By contrast, {\plant} was  designed so that the
physiological rules driving the dynamics of the system could be easily
manipulated and emergent demographic behaviours (including fitness) studied in
detail.
% From here down can be largely dropped I think
These different use-cases underpin important design differences among
the models. Whereas the primary purpose of these other models is to scale a
given physiology up to predict  broad-scale changes in the terrestrial
biosphere, {\plant} allows users to  interact with the model at multiple
levels of ecological organisation and  study the dynamics in detail. With any
given physiology, it is possible to  grow plants in isolation, in patches of
competing plants and in disturbance- driven metapopulations. {\plant} also
allows for greater focus on population dynamics, reporting not only vegetation
properties but also individual level densities, growth rates and fitness.
{\plant} thus aims to bridge the gap between existing large-scale coupled
simulation models like \textsc{ED} and the extremely simple models that are
still widely  used in theoretical ecology.
